Verifying Cache Coherence Protocols for an infinite family of cache hierarchies

### 2014-02-06 paper/Intuition.tex
this the the correct version

\section{Proof of MSI protocol being preserving Store atomicity}
\label{Sec:Safety}

In this section, we present the structure of the proof of Theorem
\fullref{storeAtomicity} for a system implementing the MSI protocol of
Figure~\ref{trans}. We reduce the whole system proof to that of a set
of invariants between any parent-child pair. These invariants in turn
depend on two smaller invariants which encapsulate the core aspects of
correctness of the MSI protocol and allow the large scope of verified
hardware refinements we present later (Section~\ref{Sec:Refinements}).

The invariants hold after any number of executions of the transactions
starting from the initial state. We use the term time $t$ to specify
the global state reached after firing $t$ transitions.

@@

\subsection{Reducing correctness to Node-level Invariants}

Store atomicity is a whole system property. Intuitively our
protocol is correct because it guarantees at most one cache may
write at a time, and that any cache permitted to read has
the ``latest'' value. We state this formally with the following 4
invariants.

\begin{inv}
\textit{latestValue}:
At any time $t$, a cache $c$ which is clean for an address $a$ will have the most
up-to-date value for that address, \ie{}
\begin{enumerate}
\item $c.\data[a] = \initData(a)$ and no store request $q'$ has been processed at
any time $t'$ such that $t' < t \wedge a = q'.\addrQ$, or
\item There is a store request $q$ that was processed at time $t_q$ such that
$t_q < t \wedge a = q.\addrQ \wedge c.\data[a] = q.\dataQ$ and no other store
request $q'$ was processed at any time $t'$ such that $t_q < t' < t \wedge a =
q'.\addrQ$.
\end{enumerate}
\label{latestValue}
\end{inv}
A clean cache is defined as follows.
\begin{defn}
\textit{clean}: A cache is said to be \clean{} for an address $a$ if and only if
the state of the cache for that address is at least \Sh{} and the directory 
state for all its children is at most \Sh.
\end{defn}

\begin{inv}
\textit{nonAncestorCompatible}: For two distinct caches $c_1$ and $c_2$ such
that neither is an ancestor of the other in the cache tree
hierarchy (where ancestor relation is the reflexive-transitive closure of the parent relation), for each
address $a$, $c_1.\state[a]$ and $c_2.\state[a]$ are \textit{compatible}.
\label{nonAncestorCompatible}
\end{inv}
Compatibility of states is defined as follows.
\begin{defn}
\textit{compatibile}: Values of states $x_1$ and $x_2$ are \textit{compatible} iff
either $x_1$ or $x_2$ is \Mo{}, then the other is \In. Note that any two states $y_1 \le x_1$ and $y_2 \le x_2$ are also compatible if $x_1, x_2$ are compatible.
\label{compatible}
\end{defn}

%\begin{inv}
%\textit{processReq}: Load and store requests can be processed at a leaf cache
%only if it is at least in the \Sh{} and \Mo{} states, respectively, for that
%address.
%\label{processReq}
%\end{inv}
%
%\begin{inv}
%\textit{noReqAgain}:
%\item For a leaf cache $c$, $c.\pos$ changes only on processing a request. In
%that case, $c.\pos$ gets incremented.
%\label{noReqAgain}
%\end{inv}

Given these invariants, we will show how Theorem~\fullref{storeAtomicity}
follows. From the transitions of Figure \ref{trans}, a store may happen only
when a leaf cache is in \Mo{} state for that
address%(Invariant~\textit{processReq}).
Leaf caches cannot be ancestors of any
other node, so by Invariant \fullref{nonAncestorCompatible}, if a leaf cache is
in \Mo{} state, all other leaf caches will be in \In{} state preventing the
possibility of processing load and store requests for the same address at the
same time. Thus, Definition~\fullref{uniqStTimes} holds.

Once a response has been produced for a request $i$, the cache increments the
counter for the request it processes next. Thus there is at most one response
to each request, satisfying Definition~\fullref{uniqRespLabels}. As leaf caches
have no children, a leaf cache in at least \Sh{} state for address $a$ is
\clean{} for that address. Since processing a load request requires the cache
to be at least in the \Sh{} state, its response, which supplies the data for
that address present in the cache, supplies the latest value by
Invariant~\fullref{latestValue}.  Thus, the final condition for
Theorem~\fullref{storeAtomicity} is also satisfied. All that remains is to show
how these invariants hold.

\subsection{Reducing to Pair-wise Invariants}
\label{pairwise}

Correctness of Invariants \fullref{localCompatibility} and \fullref{latestValue} can be
proved using the following invariants on a parent-child pair.

@@

\begin{inv}
\textit{localCompatibility}:
\begin{enumerate}
\item The state of a cache is never less than its directory entry for any of
its children.
\item Let $c_1$ and $c_2$ be distinct caches with the same parent $p$. Then,
$p.\dstate[c_1][a]$ and $p.\dstate[c_2][a]$ are \textit{compatible}.
\end{enumerate}
\label{localCompatibility}
\end{inv}

\begin{inv}
\textit{noTransitWrite}: Whenever data for an address $a$ is in transit (\ie{}
$\forall t, t_s \le t \le t_r$ where $t_s$ is the time of sending the data and
$t_r$ the time of receiving the data), no cache can process a store request for
$a$.
\label{noTransitWrite}
\end{inv}

For any two caches $c$ and $p$, such that $p$ is an ancestor of $c$ in
the cache tree hierarchy,

Invariant \fullref{conservative} and the first part of Invariant
\fullref{localCompatibility} defines a $\ge$ relation between the states of a
parent and its child for any address.  Since the ancestor relation is the
reflexive-transitive closure of the parent relation, the $\ge$ relation is
preserved between the states of two caches $p$ and $c$ where $p$ is the
ancestor of $c$.

Any two caches $c_1, c_2$ neither of which are an ancestor of the other has a
lowest common ancestor $x$ which is an ancestor of both $c_1$ and $c_2$. $x$
has two distinct children $x_1$ and $x_2$ such that they are the ancestors of
$c_1$ and $c_2$ respectively. For any address $a$, $x.\dstate[x_1][a]$ and
$x.\dstate[x_1][a]$ are compatible by the second part of
Invariant~\fullref{localCompatibility}. It follows that $x_1.\state[a]$ and
$x_2.\state[a]$ are compatible since these values are smaller than the
respective directory values.  From the previous paragraph about the relation
between the states of a cache and its ancestor, it follows that $c_1$ and $c_2$
are also compatible. This proves Invariant \fullref{nonAncestorCompatible}.

@@

If $c$ was not clean at time $t$, it must have received data from either its
parent or its child.  As can be seen in Figure~\ref{trans}, whenever data is
sent either from a parent to a child, or from a child to the parent, the sender
is \clean. By the induction hypothesis, it means that the sender has the latest
value for the address. Because of Invariant~\ref{noTransitWrite}, no store
request can be processed when the data is in transit, which means $c$ will have
the latest value for address $a$ at time $t+1$. This establishes Invariant
\fullref{latestValue}.

\subsection{Proving parent-child hierarchy properties}

%ndave: This needs some repointing

We present two basic invariants that the MSI protocol of Figure~\ref{trans}
obeys. In fact the request/response message semantics was designed to satisfy
the following two invariants. We believe that most cache-coherence protocols
would end up obeying these two invariants. With two invariants we can
prove all the pair-wise properties of Section~\ref{pairwise}.

@@

The MSI-based protocol in Figure~\ref{trans} is such that any change in either
the state of a cache or its directory entry takes place only on sending or
receiving a response. The parent upgrades the corresponding directory entry on
sending a response to a child and the child downgrades its state on
sending a response to the parent.

To prove Invariant \fullref{conservative} at some time $t$, consider the case
when a child cache $c$ has sent to or received from the parent $p$ some message
before $t$. Consider the last such message $m$ sent by or received at $c$ at
$t'$. Let's say it is sent by $c$ at $t'$, not received. Then $p$ can not have
received any message from $c$ after sending $m$ and before $t$ in order to not
violate Invariant \fullref{respFIFO}.  Instead, if $c$ receives $m$ at $t'$,
$p$ will not receive any message after sending $m$ and before $t$ in order to
not violate Invariant \fullref{noCross}. Thus, $p$ will not receive any message
from $c$ after sending or receiving $m$ and before $t$.  Let $t''$ be the time
at which $p$ has sent or received $m$. Whether $p$ sends or receives $m$,
$p.\dstate[c][a]$ at $t''+1$ will be equal to $c.\state[a]$ at $t'+1$ because
the response messages sync up the next-state according to the transitions in
Figure \ref{trans}. Since $p$ can not receive any message between $t''+1$ and
$t$, it can only upgrade $p.\dstate[c][a]$. $c.\state[a]$ does not change
between $t'+1$ and $t$ as it never changes state. Therefore, Invariant
\fullref{conservative} holds at $t$.

If no message has been sent or received by $c$ before $t$, then $p$ wouldn't
have received any message sent by $c$ before $t$. So $p.\dstate[c][a]$ at time
$t \ge p.\dstate[c][a]$ at time $0$ . Since the initial values of the directory
and the child's state are both set to \In{} Invariant \textit{conservative}
holds.

To prove Invariant \fullref{localCompatibility}, it is enough to show that a
$c.\state[a]$ is upgraded on receiving a response for $a$ from $c$'s parent $p$
and $p.\dstate[c][a]$ is downgraded on receiving a response for $a$ from $c$.
If this is true, then Invariant \fullref{localCompatibility} will not be
violated on receiving a response. The transitions of Figure \ref{trans} ensures
that Invariant \fullref{localCompatibility} is not violated when sending a
response. Since whenever the state or directory changes, the appropriate
response is sent or received, Invariant \fullref{localCompatibility} holds.

If the value of the state of a child at the time of sending a response is the
same as the value of the directory for that child at the parent at the time of
receiving the response, it follows straightaway that the parent downgrades its
directory on receiving a response from the child. This is because a child
downgrades on sending a response and the next-states are synced on receiving
the response.  This proof uses exactly the same arguments presented above to
prove Invariant \fullref{conservative} -- we prove by considering the last
message sent or received at the child before $t$ (and the case when no message
has been sent or received). Similarly, we can show that a child upgrades its
state on receiving a response, satisfying the conditions in the previous paragraph.

Invariant \fullref{noTransitWrite} can be proved as follows. Consider a
response sent from a parent $p$ at time $t$ received by its child $c$ at time
$t'$. No other response can be received by the parent at time $t_p$ such that
$t+1 \le t_p \le t'$ in order to not violate Invariant \fullref{noCross}, so
$p.\dstate[c][a]$ at $t+1 \le p.\dstate[c][a]$ at $t'$, since the parent can
only upgrade its directory for $c$. Similarly, no response can be sent by the
child at time $t_c$ such that $t \le t_c \le t'$ for the same reason and hence
$c.\state[a]$ at $t \le c.\state[a]$ at $t'$ since the child can only downgrade
its state.  This means that during $t_p$ such that $t+1 \le t_p \le t'$,
$p.\dstate[c][a] > \In$ and (hence $p.\state[a] > \In$ by the first part of
Invariant \fullref{localCompatibility}) since the parent upgraded its directory
for $c$ at $t$.  Similarly, during $t_c$ such that $t \le t_c \le t'$,
$c.\state[a] < \Mo$ since the child upgrades its state at $t'$. We have already
shown that (in the proof of Invariant \fullref{latestValue}) if any cache is
not in \In{} state, no non-descendent cache can process a store request. In
this case, during the times between $t$ and $t'$, $p$ is not in \In{} state.
Among $p$'s directory entries, all of them save the entry for $c$ is not in
\Mo{} state, so none of the descendents of $p$'s children, save $c$'s
descendents can process a store request. During the times between $t$ and $t'$,
$c$'s state is $< \Mo{}$, so neither $c$ nor its descendents can process a store
request, proving Invariant \fullref{noTransitWrite}.

>>>>

\section{Proof of MSI protocol being preserving Store atomicity}
\label{safety}

In this section, we present the structure of the proof of Theorem
\fullref{storeAtomicity} for a system implementing the MSI protocol of
Figure~\ref{trans}.

We first reduce the proof of store atomicity to 4 invariants, presented in
Section \ref{firstLevel}. These invariants in turn can be reduced to 4 more
invariants presented in Section \ref{secondLevel}. The reason for choosing the 8
 invariants is because they are not specific to MSI protocol
alone; any other protocol which obeys these invariants obeys Theorem
\fullref{storeAtomicity}. We finally present the key insight which we
discovered for proving the second-level invariants in the form of 3 invariants.
We believe that these properties are common in most, if not all
invalidation-based protocols which do not perform silent drops (\ie{} protocols
in which a cache that downgrades its state informs some other cache about the
downgrade). We informally show why this is the case using the example of MESI,
MOSI and MOESI in Section \ref{mosi}.

%We use the terms write to mean handling a store request, \ie{} executing the
%StoreReq transition and read to mean handling a load request, \ie{} executing
%the LoadReq transition.

These invariants hold after any number of executions of the transactions starting
from the initial state. We use the term time $t$ to specify the global state
reached after firing $t$ transitions.

@@

\subsection{First level invariants}
Intuitively our protocol is correct because it guarantees that at most one
cache can handle a store request at a time, and that any cache that attempts to
handle a load request has the ``latest'' value. We can state this formally with
the following 4 invariants.

\begin{inv}
\textit{latestValue}:
At any time $t$, a cache $c$ which is clean for an address $a$ will have the most
up-to-date value for that address, \ie{}
\begin{enumerate}
\item $c.\data[a] = \initData(a)$ and no store request $q'$ has been processed at
any time $t'$ such that $t' < t \wedge a = q'.\addrQ$, or
\item There is a store request $q$ that was processed at time $t_q$ such that
$t_q < t \wedge a = q.\addrQ \wedge c.\data[a] = q.\dataQ$ and no other store
request $q'$ was processed at any time $t'$ such that $t_q < t' < t \wedge a =
q'.\addrQ$.
\end{enumerate}
\label{latestValue}
\end{inv}
A clean cache is defined as follows.
\begin{defn}
\textit{clean}: A cache is said to be \clean{} for an address $a$ if and only if
the state of the cache for that address is at least \Sh{} and the directory 
state for all its children is at most \Sh.
\label{clean}
\end{defn}

\begin{inv}
\textit{nonAncestorCompatible}: For two distinct caches $c_1$ and $c_2$ such
that neither is an ancestor of the other in the cache tree hierarchy (where
ancestor relation is the reflexive-transitive closure of the parent relation),
for each address $a$, $c_1.\state[a]$ and $c_2.\state[a]$ are
\textit{compatible}.
\label{nonAncestorCompatible}
\end{inv}
Compatibility of states is defined as follows.
\begin{defn}
\textit{compatibile}: Values of states $x_1$ and $x_2$ are \textit{compatible}
iff either $x_1$ or $x_2$ is \Mo{}, then the other is \In.
\label{compatible}
\end{defn}

It is easy to see that any two states $y_1 \le x_1$ and $y_2 \le x_2$ are also
compatible if $x_1, x_2$ are compatible.

\begin{inv}
\textit{processReq}: Load and store requests can be processed only at a leaf
node which is at least in the \Sh{} and \Mo{} states, respectively, for that
address.
\label{processReq}
\end{inv}

\begin{inv}
\textit{noReqAgain}:
\item For a leaf cache $c$, $c.\pos$ changes only on processing a request. In
that case, $c.\pos$ gets incremented.
\label{noReqAgain}
\end{inv}

Given these invariants, we will show how Theorem~\fullref{storeAtomicity}
follows.
%From the transitions of Figure \ref{trans}, a store may happen only
%when a leaf cache is in \Mo{} state for that
%address (Invariant~\textit{processReq}).
Leaf caches cannot be ancestors of any other node, so by Invariant
\fullref{nonAncestorCompatible}, if a leaf cache is in \Mo{} state, all other
leaf caches will be in \In{} state preventing the possibility of processing
load and store requests (because of Invariant \ref{processReq}) for the same
address at the same time at different leaf caches. This satisfies
Definition~\fullref{uniqStTimes}.

It is easy to see that Definition \fullref{uniqRespLabels}, Definition
\fullref{localOrdering} and Definition \fullref{allPrevious} directly follows
from Invariant \fullref{noReqAgain}.  As leaf caches have no children, one in
at least \Sh{} state for address $a$ is \clean{} for that address. Since
processing a load request requires the cache to be at least in the \Sh{} state,
its response, which supplies the data for that address present in the cache,
supplies the latest value by Invariant~\fullref{latestValue}. Thus, all the
conditions for Theorem~\fullref{storeAtomicity} is satisfied given Invariants
\ref{latestValue} to \ref{noReqAgain}.

\subsection{Second-level Invariants}
\label{second-level}

While Invariant \fullref{processReq} and \fullref{noReqAgain} can be seen
directly from the MSI transitions, correctness of Invariants
\fullref{localCompatibility} and \fullref{latestValue} require further
decomposition. We present 4 more invariants for proving those two invariants.

@@

\begin{inv}
\textit{localCompatibility1}:
The state of a cache is never less than its directory entry for any of
its children.
\label{localCompatibility1}
\end{inv}
\begin{inv}
\textit{localCompatibility2}: Let $c_1$ and $c_2$ be distinct caches with the
same parent $p$. Then, $c_1.\state[a]$ and $c_2.\state[a]$ are
\textit{compatible}.
\label{localCompatibility2}
\end{inv}

\begin{inv}
\textit{noTransitWrite}: Whenever data for an address $a$ is in transit (\ie{}
$\forall t, t_s \le t \le t_r$ where $t_s$ is the time of sending the data and
$t_r$ the time of receiving the data), no cache can process a store request for
$a$, and the data must be sent from a cache which is \textit{clean}.
\label{noTransitWrite}
\end{inv}

Invariants \fullref{conservative} and \fullref{localCompatibility1} defines a
$\ge$ relation between the states of a parent and its child for any address.
Since the ancestor relation is the reflexive-transitive closure of the parent
relation, the $\ge$ relation is preserved between the states of two caches $p$
and $c$ where $p$ is the ancestor of $c$.

Any two caches $c_1, c_2$ neither of which are an ancestor of the other has a
lowest common ancestor $x$ which is an ancestor of both $c_1$ and $c_2$. $x$
has two distinct children $x_1$ and $x_2$ such that they are the ancestors of
$c_1$ and $c_2$ respectively. For any address $a$, $x_1.\state[a]$ and
$x_2.\state[a]$ are compatible by the second part of
Invariant~\fullref{localCompatibility}. It follows that $x_1.\state[a]$ and
$x_2.\state[a]$ are compatible since these values are smaller than the
respective directory values. From the previous paragraph about the relation
between the states of a cache and its ancestor, it follows that $c_1$ and $c_2$
are also compatible. This proves Invariant \fullref{nonAncestorCompatible}.

@@

If $c$ was not clean at time $t$, it must have received data from either its
parent or its child. And the sender is \clean{} by Invariant
\ref{noTransitWrite}. By the induction hypothesis, it means that the sender has
the latest value for the address. Again, because of
Invariant~\ref{noTransitWrite}, no store request can be processed when the data
is in transit, which means $c$ will have the latest value for address $a$ at
time $t+1$. This establishes Invariant \fullref{latestValue}.

\subsection{Key insight for proving second-level invariants}

%ndave: This needs some repointing
We present what we believe to be the key insight for proving any cache coherence protocol.
In fact the request/response message semantics for the MSI protocol was designed to satisfy
the following three invariants.

\begin{inv}
\textit{stateChange}:
\begin{enumerate}
\item A cache downgrades its state for an address iff it sends a response for that address
\item A parent upgrades the directory entry for an address for one of its
children iff it sends a response for that child and address
\item The state or directory entry changes for an address only on sending and
receiving an response for that address, and they change to the \myto-field of the response.
\end{enumerate}
\label{stateChange}
\end{inv}

Invariant \fullref{stateChange} ensures that a response message acts as a sync
between a parent's directory entry for a child and the state of the child, as
both the sender and the receiver change their respective directory entry or
state to the \myto-field of the response.

@@

\begin{figure}
\centering
\includegraphics[scale=.5]{resps}
\caption{An example showing a sequence of responses ping ponging between a child and its parent}
\label{resps}
\end{figure}

Figure \ref{resps} shows an example of the effect of Invariants
\fullref{stateChange}, \fullref{noCross} and \fullref{respFIFO}. The directory
entry of the parent or the state of the child changes on sending and receiving
a response and the new state is showed at the points of sending or receiving
responses.
Both the cache and its parent follow the same sequence state and directory
entry changes; just that the receiver always lags slightly behind the sender.
The receiver can not send any response before receiving because of
\fullref{noCross}, thereby eliminating any scenarios where the lagging makes
reasoning complicated.

Using these three invariants (Invariant \ref{stateChange} to \ref{respFIFO}, it
is easy to see that Invariant \fullref{conservative} holds. Intuitively, going
back to Figure \ref{resps}, we can see that at all times, either there is no
response in flight, or only responses from parent in flight, or responses from
child in flight. A response from child downgrades the state and a response from
parent upgrades the directory. Since the other side simply lags behind,
Invariant \fullref{conservative} holds.

Proving the other 3 invariants in Section \ref{second-level} requires using
more properties of the MSI transistions of Figure \ref{trans}, though it is
straightforward.

%To prove Invariant \fullref{conservative} at some time $t$, consider the case
%when a child cache $c$ has sent to or received from the parent $p$ some message
%before $t$. Consider the last such message $m$ sent by or received at $c$ at
%$t'$. Let's say it is sent by $c$ at $t'$, not received. Then $p$ can not have
%received any message from $c$ after sending $m$ and before $t$ in order to not
%violate Invariant \fullref{respFIFO}.  Instead, if $c$ receives $m$ at $t'$,
%$p$ will not receive any message after sending $m$ and before $t$ in order to
%not violate Invariant \fullref{noCross}. Thus, $p$ will not receive any message
%from $c$ after sending or receiving $m$ and before $t$.  Let $t''$ be the time
%at which $p$ has sent or received $m$. Whether $p$ sends or receives $m$,
%$p.\dstate[c][a]$ at $t''+1$ will be equal to $c.\state[a]$ at $t'+1$ because
%the response messages sync up the next-state according to the transitions in
%Figure \ref{trans}. Since $p$ can not receive any message between $t''+1$ and
%$t$, it can only upgrade $p.\dstate[c][a]$. $c.\state[a]$ does not change
%between $t'+1$ and $t$ as it never changes state. Therefore, Invariant
%\fullref{conservative} holds at $t$.
%If no message has been sent or received by $c$ before $t$, then $p$ wouldn't
%have received any message sent by $c$ before $t$. So $p.\dstate[c][a]$ at time
%$t \ge p.\dstate[c][a]$ at time $0$ . Since the initial values of the directory
%and the child's state are both set to \In{} Invariant \textit{conservative}
%holds.
%To prove Invariant \fullref{localCompatibility}, it is enough to show that a
%$c.\state[a]$ is upgraded on receiving a response for $a$ from $c$'s parent $p$
%and $p.\dstate[c][a]$ is downgraded on receiving a response for $a$ from $c$.
%If this is true, then Invariant \fullref{localCompatibility} will not be
%violated on receiving a response. The transitions of Figure \ref{trans} ensures
%that Invariant \fullref{localCompatibility} is not violated when sending a
%response. Since whenever the state or directory changes, the appropriate
%response is sent or received, Invariant \fullref{localCompatibility} holds.
%If the value of the state of a child at the time of sending a response is the
%same as the value of the directory for that child at the parent at the time of
%receiving the response, it follows straightaway that the parent downgrades its
%directory on receiving a response from the child. This is because a child
%downgrades on sending a response and the next-states are synced on receiving
%the response.  This proof uses exactly the same arguments presented above to
%prove Invariant \fullref{conservative} -- we prove by considering the last
%message sent or received at the child before $t$ (and the case when no message
%has been sent or received). Similarly, we can show that a child upgrades its
%state on receiving a response, satisfying the conditions in the previous paragraph.
%Invariant \fullref{noTransitWrite} can be proved as follows. Consider a
%response sent from a parent $p$ at time $t$ received by its child $c$ at time
%$t'$. No other response can be received by the parent at time $t_p$ such that
%$t+1 \le t_p \le t'$ in order to not violate Invariant \fullref{noCross}, so
%$p.\dstate[c][a]$ at $t+1 \le p.\dstate[c][a]$ at $t'$, since the parent can
%only upgrade its directory for $c$. Similarly, no response can be sent by the
%child at time $t_c$ such that $t \le t_c \le t'$ for the same reason and hence
%$c.\state[a]$ at $t \le c.\state[a]$ at $t'$ since the child can only downgrade
%its state.  This means that during $t_p$ such that $t+1 \le t_p \le t'$,
%$p.\dstate[c][a] > \In$ and (hence $p.\state[a] > \In$ by the first part of
%Invariant \fullref{localCompatibility}) since the parent upgraded its directory
%for $c$ at $t$.  Similarly, during $t_c$ such that $t \le t_c \le t'$,
%$c.\state[a] < \Mo$ since the child upgrades its state at $t'$. We have already
%shown that (in the proof of Invariant \fullref{latestValue}) if any cache is
%not in \In{} state, no non-descendent cache can process a store request. In
%this case, during the times between $t$ and $t'$, $p$ is not in \In{} state.
%Among $p$'s directory entries, all of them save the entry for $c$ is not in
%\Mo{} state, so none of the descendents of $p$'s children, save $c$'s
%descendents can process a store request. During the times between $t$ and $t'$,
%$c$'s state is $< \Mo{}$, so neither $c$ nor its descendents can process a store
%request, proving Invariant \fullref{noTransitWrite}.
